Anomaly detection with kernel density estimation on manifolds

### 2022-02-25 paper/paper.Rmd
Update Swissroll plot, table full.width=F

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  messages = FALSE,
  warning = FALSE,
  eval = TRUE
  # include = FALSE
)
options(tinytex.verbose = TRUE)
library(tidyverse)
library(dimRed)
library(reticulate)
library(here)
library(viridis)
library(hdrcde)
library(igraph)
library(matrixcalc)
library(akima)
library(car)
library(ggforce)
library(ks)
library(patchwork)
# library(copula)
library(plotly)
library(kableExtra)
Jmisc::sourceAll(here::here("R/sources"))
# set.seed(123)
```

@@

<!-- ### Swiss roll mapping -->

@@

```{r sroutliers, fig.align = 'center', out.width = "95%", fig.cap="Highest density region plots of four manifold learning embeddings of the swiss roll data. Colors are indicating densities from top: true densities from the Gaussian mixture model; middle: KDE with Riemannian matrix as variable bandwidth; and bottom: KDE with fixed bandwidth. Variable KDE preforms better in finding kernel structures with ISOMAP and LLE, and in locating outliers with t-SNE. Both methods are not detecting outliers accurately when using UMAP.", echo = FALSE}
knitr::include_graphics("figures/Swiss Roll_outliers_comparison_4ml_3cases_riem001.png")
```

@@

```{r srcors, echo=FALSE, message=FALSE, eval=F}
load("figures/CorrelationTable_Swiss Roll_4ml_riem01.rda")
col.max <- apply(cors, 2, max)
max.match <- rep(col.max, each = 2) == cors
data.frame(cors) %>% 
  mutate_if(is.numeric, format, digits = 3, nsmall = 3) %>% 
  kableExtra::kbl(caption = "Correlation between true density and estimated density for four manifold learning embeddings.", booktabs = TRUE, digits = 3, escape = FALSE) %>%
  kable_styling(latex_options = "scale_down") %>%
  # kable_paper(full_width = TRUE) %>%
  column_spec(2, bold = max.match[,1]) %>%
  column_spec(3, bold = max.match[,2]) %>%
  column_spec(4, bold = max.match[,3]) %>%
  column_spec(5, bold = max.match[,4]) %>% 
  column_spec(1, width = "6cm")
```

@@

```{r tpcors, echo=FALSE, message=FALSE, eval=T}
load("figures/CorrelationTable_twinpeak_4ml_riem01.rda")
col.max <- apply(cors, 2, max)
max.match <- rep(col.max, each = 2) == cors
data.frame(cors) %>% 
  mutate_if(is.numeric, format, digits = 3, nsmall = 3) %>% 
  kableExtra::kbl(caption = "Correlation between true density and estimated density for four manifold learning embeddings.", booktabs = TRUE, digits = 3, escape = FALSE) %>%
  kable_styling(latex_options = "scale_down") %>%
  # kable_paper(full_width = TRUE) %>%
  column_spec(2, bold = max.match[,1]) %>%
  column_spec(3, bold = max.match[,2]) %>%
  column_spec(4, bold = max.match[,3]) %>%
  column_spec(5, bold = max.match[,4]) %>% 
  column_spec(1, width = "6cm")
```

@@

```{r fivedcors, echo=FALSE, message=FALSE, eval=T}
load("figures/CorrelationTable_fived_4ml_riem20.rda")
col.max <- apply(cors, 2, max)
max.match <- rep(col.max, each = 2) == cors
data.frame(cors) %>% 
  mutate_if(is.numeric, format, digits = 3, nsmall = 3) %>% 
  kableExtra::kbl(caption = "Correlation between true density and estimated density for four manifold learning embeddings.", booktabs = TRUE, digits = 3, escape = FALSE) %>%
  kable_styling(latex_options = "scale_down") %>%
  # kable_paper(full_width = TRUE) %>%
  column_spec(2, bold = max.match[,1]) %>%
  column_spec(3, bold = max.match[,2]) %>%
  column_spec(4, bold = max.match[,3]) %>%
  # column_spec(5, bold = max.match[,4]) %>% 
  column_spec(1, width = "6cm")
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  messages = FALSE,
  warning = FALSE,
  eval = TRUE
  # include = FALSE
)
options(tinytex.verbose = TRUE)
library(tidyverse)
library(dimRed)
library(reticulate)
library(here)
library(viridis)
library(hdrcde)
library(igraph)
library(matrixcalc)
library(akima)
library(car)
library(ggforce)
library(ks)
library(patchwork)
# library(copula)
library(plotly)
library(kableExtra)
Jmisc::sourceAll(here::here("R/sources"))
# set.seed(1234)
```

@@

### Swiss roll mapping

@@

```{r sroutliers, fig.align = 'center', out.width = "95%", fig.cap="Highest density region plots of four manifold learning embeddings of the swiss roll data. Colors are indicating densities from top: true densities from the Gaussian mixture model; middle: KDE with Riemannian matrix as variable bandwidth; and bottom: KDE with fixed bandwidth. Variable KDE preforms better in finding kernel structures with ISOMAP and LLE, and in locating outliers with t-SNE. Both methods are not detecting outliers accurately when using UMAP.", echo = FALSE}
knitr::include_graphics("figures/Swiss Roll_outliers_comparison_4ml_3cases_riem0_08.png")
```

@@

```{r srcors, echo=FALSE, message=FALSE, eval=T}
load("figures/CorrelationTable_Swiss Roll_4ml_riem0_08.rda")
col.max <- apply(cors, 2, max)
max.match <- rep(col.max, each = 2) == cors
data.frame(cors) %>% 
  mutate_if(is.numeric, format, digits = 3, nsmall = 3) %>% 
  kableExtra::kbl(caption = "Correlation between true density and estimated density for four manifold learning embeddings.", booktabs = TRUE, digits = 3, escape = FALSE) %>%
  kable_styling(latex_options = "scale_down") %>%
  # kable_paper(full_width = FALSE) %>%
  # column_spec(1, width = "6cm") %>% 
  column_spec(2, bold = max.match[,1]) %>%
  column_spec(3, bold = max.match[,2]) %>%
  column_spec(4, bold = max.match[,3]) %>%
  column_spec(5, bold = max.match[,4]) %>%
  column_spec(6, bold = max.match[,5])
```

@@

```{r tpcors, echo=FALSE, message=FALSE, eval=T}
load("figures/CorrelationTable_twinpeak_4ml_riem01.rda")
col.max <- apply(cors, 2, max)
max.match <- rep(col.max, each = 2) == cors
data.frame(cors) %>% 
  mutate_if(is.numeric, format, digits = 3, nsmall = 3) %>% 
  kableExtra::kbl(caption = "Correlation between true density and estimated density for four manifold learning embeddings.", booktabs = TRUE, digits = 3, escape = FALSE) %>%
  # kable_styling(latex_options = "scale_down") %>%
  kable_paper(full_width = FALSE) %>%
  # column_spec(1, width = "6cm") %>% 
  column_spec(2, bold = max.match[,1]) %>%
  column_spec(3, bold = max.match[,2]) %>%
  column_spec(4, bold = max.match[,3]) %>%
  column_spec(5, bold = max.match[,4]) #%>%
  # column_spec(6, bold = max.match[,5])
```

@@

```{r fivedcors, echo=FALSE, message=FALSE, eval=T}
load("figures/CorrelationTable_fived_4ml_riem20.rda")
col.max <- apply(cors, 2, max)
max.match <- rep(col.max, each = 2) == cors
data.frame(cors) %>% 
  mutate_if(is.numeric, format, digits = 3, nsmall = 3) %>% 
  kableExtra::kbl(caption = "Correlation between true density and estimated density for four manifold learning embeddings.", booktabs = TRUE, digits = 3, escape = FALSE) %>%
  # kable_styling(latex_options = "scale_down") %>% # automatically adjust the table to page width
  kable_paper(full_width = FALSE) %>%
  # column_spec(1, width = "6cm") %>% 
  column_spec(2, bold = max.match[,1]) %>%
  column_spec(3, bold = max.match[,2]) %>%
  column_spec(4, bold = max.match[,3]) #%>%
  # column_spec(5, bold = max.match[,4]) 
```

### 2022-02-25 paper/paper.tex
Update Swissroll plot, table full.width=F

@@

{\centering \includegraphics[width=0.95\linewidth]{figures/Swiss Roll_outliers_comparison_4ml_3cases_riem001}

@@

@@

\caption{\label{tab:tpcors}Correlation between true density and estimated density for four manifold learning embeddings.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6cm}>{}l>{}l>{}l>{}l}
\toprule
  & ISOMAP & LLE & t.SNE & UMAP\\
\midrule
Variable bandwidth & \textbf{0.940} & \textbf{0.407} & 0.520 & 0.259\\
Fixed bandwidth & 0.626 & 0.399 & \textbf{0.660} & \textbf{0.655}\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:fivedcors}Correlation between true density and estimated density for four manifold learning embeddings.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6cm}>{}l>{}l>{}l}
\toprule
  & ISOMAP & LLE & UMAP\\
\midrule
Variable bandwidth & \textbf{0.921} & \textbf{0.981} & \textbf{-0.130}\\
Fixed bandwidth & 0.806 & 0.940 & -0.341\\
\bottomrule
\end{tabular}}
\end{table}

>>>>

\hypertarget{swiss-roll-mapping}{%
\subsubsection{Swiss roll mapping}\label{swiss-roll-mapping}}

@@

{\centering \includegraphics[width=0.95\linewidth]{figures/Swiss Roll_outliers_comparison_4ml_3cases_riem0_08}

@@

\caption{\label{tab:srcors}Correlation between true density and estimated density for four manifold learning embeddings.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{l>{}l>{}l>{}l>{}l>{}l}
\toprule
  & ISOMAP & LLE & Laplacian.Eigenmaps & t.SNE & UMAP\\
\midrule
Variable bandwidth & 0.0696 & \textbf{0.400} & -0.2357 & 0.023 & \textbf{0.0138}\\
Fixed bandwidth & \textbf{0.2798} & 0.351 & \textbf{0.0141} & \textbf{0.367} & -0.0110\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:tpcors}Correlation between true density and estimated density for four manifold learning embeddings.}
\centering
\begin{tabular}[t]{l>{}l>{}l>{}l>{}l}
\toprule
  & ISOMAP & LLE & t.SNE & UMAP\\
\midrule
Variable bandwidth & \textbf{0.940} & \textbf{0.407} & 0.520 & 0.259\\
Fixed bandwidth & 0.626 & 0.399 & \textbf{0.660} & \textbf{0.655}\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:fivedcors}Correlation between true density and estimated density for four manifold learning embeddings.}
\centering
\begin{tabular}[t]{l>{}l>{}l>{}l}
\toprule
  & ISOMAP & LLE & UMAP\\
\midrule
Variable bandwidth & \textbf{0.921} & \textbf{0.981} & \textbf{-0.130}\\
Fixed bandwidth & 0.806 & 0.940 & -0.341\\
\bottomrule
\end{tabular}
\end{table}
